An Accurate Model for SARS-CoV-2 Pooled RT-PCR Test Errors

### 2021-09-10 paper.tex
merge

\section*{Introduction}
Reverse transcription polymerase chain reaction (RT-PCR) testing is a
key component in breaking transmission chains and mitigating the
COVID-19 pandemic. As such, the need for large-scale testing has
resulted in the application of techniques to increase the throughput
of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30, BayesianDorfman,
  MatrixPooling, LionDorfman, CherifReview}. These techniques, often
termed \emph{pooling}, or group testing, have originated in the 1940's
in the seminal work of Dorfman and his pooling technique
\cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one
selects $N$ individuals and performs a single RT-PCR test on their
combined (\emph{pooled}) samples. If the pooled test yields a positive
result, then each individual is retested separately; otherwise,
everyone is declared negative. The throughput efficiency of Dorfman
pooling has been demonstrated empirically \cite{DorfmanYuvalDor} and
its error rates thoroguhly investigated \cite{Kim, Simplistic1,
  OptimalDorfmanPool}. Other commonly used techniques include
\emph{recursive} pooling \cite{Kim,RecursiveSevenFold}, wherein if the
first pooled test is positive, the pool is split into two and the
process repeats. Otherwise, all pool members are declared negative.
In \emph{Matrix} pooling \cite{MatrixPooling} a population of size
$N=mn$ is arranged in a matrix of dimensions $m\times n$ matrix. Each
row and column are then pooled and individuals in the intersection of
positive rows and columns are tested separately.

@@

Several consequences arise from our analysis. Estimation of optimal
pool size (for Dorfman pooling), which relies on previously made
asumptions should be revisited \cite{OptimalDorfmanPool}. The error
rates of any pooling scheme in use should be reesetimated
\cite{BayesianDorfman, Kim}. And, finally, new methods should be
developed in order to take advantage of our novel model \cite{DOPE}.

@@

\section*{Discussion}
In this study, we developed a more realistic and novel probabilistic
model for the outcomes of pooled PCR tests, parametrized for
SARS-CoV-2. Contrary to the common assumption, we assume, based on
data (Table~\ref{table}), that multiple infected individuals increase
the likelihood of a positive pooled test result. A direct consequence
of our model is that low values of infection prevalence increase the
false-negative rates of Dorfman pooling. Importantly, our results
remain qualitatively similar under varying parameter values, in the
observed ranges \cite{KitComparison,EstimatingRatesKucrika,
  EstimatingRatesLourenco, InterpretingCOVID19Test} (Figure
\ref{fig1}). Therefore, our results give rise to a conflict: low
infection prevalence leads to high efficiency of Dorfman pooling
\cite{DorfmanYuvalDor}, while also increasing false-negative rates.

@@

Furthermore, an especially important consideration in designing
pooling schemes is explicitly taking the intrinsic RT-PCR error rates
into account. As we have demonstrated above, analyzing the expected
error rates of pooling schemes under a realistic error
model is readily achievable. Designing pooling schemes which optimize
testing while accounting for error rates is, however, considerably more involved.
Nonetheless, various algorithms explicitly incorporate error
rates into their pooling optimization processes \cite{BayesianDorfman, Kim, OptimalDorfmanPool}, including a
recent study by our group which employed the herein described error
model \cite{DOPE}.  In our proposed algorithm, we have used a Bayesian
experimental design criterion to create an algorithm which iteratively
performs pooling, updating its information based on previous
results. In doing so, our pooling scheme shows potential in terms of
both reducing false-negative and false-positive rates, as well as increasing efficiency,
although it has not yet been implemented in real-world settings.

>>>>

\section*{Introduction}
Reverse transcription polymerase chain reaction (RT-PCR) testing is a
key component in breaking transmission chains and mitigating the
COVID-19 pandemic. As such, the need for large-scale testing has
resulted in the application of techniques to increase the throughput
of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30, BayesianDorfman,
  MatrixPooling, LionDorfman, CherifReview}. These techniques, often
termed \emph{pooling}, or group testing, have originated in the 1940's
in the seminal work of Dorfman and his pooling technique
\cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one
selects $N$ individuals and performs a single RT-PCR test on their
combined (\emph{pooled}) samples. If the pooled test yields a positive
result, then each individual is retested separately; otherwise,
everyone is declared negative. The throughput efficiency of Dorfman
pooling has been demonstrated empirically \cite{DorfmanYuvalDor} and
its error rates thoroughly investigated \cite{Kim, Simplistic1,
  OptimalDorfmanPool}.

The study and practice of pooling has advanced considerably since Dorfman's original paper, with 
several other pooling techniques being applied for SARS-CoV-2.
For example, in \emph{recursive} pooling \cite{Kim,RecursiveSevenFold}, if the
first pooled test is positive, the pool is split into two and the
process repeats. Otherwise, all pool members are declared negative.
In \emph{Matrix} pooling \cite{MatrixPooling} a population of size
$N=mn$ is arranged in a matrix of dimensions $m\times n$. Each
row and column are then pooled and individuals in the intersection of
positive rows and columns are tested separately. Other, more sophisticated
pooling schemes exist \cite{CompressedPooling}, but studies investigating the 
underlying assumptions of pooling are surprisingly lacking.

@@

Several consequences arise from our analysis: Previously estimated optimal Dorfman
pool size should be revisited, as it is influenced by the prevalence \cite{OptimalDorfmanPool}. The error
rates of many pooling schemes in use should be reesetimated
\cite{BayesianDorfman, Kim}. And, finally, new methods should be
developed in order to take advantage of our novel model \cite{DOPE}.

@@

\section*{Discussion}
In this study, we developed a more realistic and novel probabilistic
model for the outcomes of pooled RT-PCR tests, parametrized for
SARS-CoV-2. Contrary to the common assumption, we assume, based on
data (Table~\ref{table}), that multiple infected individuals increase
the likelihood of a positive pooled test result. A direct consequence
of our model is that low values of infection prevalence increase the
false-negative rates of Dorfman pooling. Importantly, our results
remain qualitatively similar under varying parameter values, in the
observed ranges \cite{KitComparison,EstimatingRatesKucrika,
  EstimatingRatesLourenco, InterpretingCOVID19Test} (Figure
\ref{fig1}). Therefore, our results give rise to a conflict: low
infection prevalence leads to high efficiency of Dorfman pooling
\cite{DorfmanYuvalDor}, while also increasing false-negative rates.

@@

Furthermore, an especially important consideration in designing
pooling schemes is explicitly taking the intrinsic RT-PCR error rates
into account. As we have demonstrated above, analyzing the expected
error rates of pooling schemes under a realistic error
model is readily achievable. Designing pooling schemes which optimize
testing while accounting for error rates is, however, considerably more involved.
Nonetheless, various algorithms explicitly incorporate error
rates into their pooling optimization processes \cite{Kim, OptimalDorfmanPool}, including a
recent study by our group which employed the herein described error
model \cite{DOPE}.  In our proposed algorithm, we have used an
experimental design criterion to create an algorithm which iteratively
performs pooling, updating its information based on previous
results. In doing so, our pooling scheme shows potential in terms of
both reducing false-negative and false-positive rates, as well as increasing efficiency,
although it has not yet been implemented in real-world settings.
